---
title: "Reproducible, collaborative workflow with Github and R Markdown"
author: "Julie Lowndes"
date: "July 13, 2016"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmltools)
```

# Objectives



```{r, message=F, warning=F, results='hide'}
## load libraries
library(dplyr) # you need to have all libraries loaded within the Rmd file
library(ggplot2) #install.packages('ggplot2')
library(gapminder)

# get range of available data
summary(gapminder)
```

### Scatterplot

```{r, message=F, warning=F}
# setup dataframe
g = gapminder %>%
  filter(year==2007) %>%   # most recent year 
  mutate(pop_m = pop/1e6)  # population, millions

# plot scatterplot of most recent year 
s = ggplot(g, aes(x=gdpPercap, y=lifeExp)) +
  geom_point()
s

# add aesthetic of size by population
s = s + 
  aes(size=pop_m)
s

# add aesthetic of color by continent
s = s + 
  aes(color=continent)
s

# add title, update axes labels
s = s + 
  ggtitle('Health & Wealth of Nations for 2007') +
  xlab('GDP per capita ($/year)') +
  ylab('Life expectancy (years)')
s

# label legend
s = s + 
  scale_colour_discrete(name='Continent') +
  scale_size_continuous(name='Population (M)')
s
```

**Your Turn**: Make a similar plot but for `gdpPercap`. Be sure to update the plot's aesthetic, axis label and title accordingly.

### Boxplot

```{r}
# boxplot by continent
b = ggplot(g, aes(x=continent, y=lifeExp)) +
  geom_boxplot()
b

# match color to continents, like scatterplot
b = b +
  aes(fill=continent)
b

# drop legend, add title, update axes labels
b = b +
  theme(legend.position='none') +
  ggtitle('Life Expectancy by Continent for 2007') +
  xlab('Continent') +
  ylab('Life expectancy (years)')
b
```

## Interactive: `plotly`

[ggplot2 | plotly](https://plot.ly/ggplot2/)

```{r, eval = FALSE, message=F, warning=F}
library(plotly) # install.packages('plotly')

# scatterplot (Note: key=country shows up on rollover)
s = ggplot(g, aes(x=gdpPercap, y=lifeExp, key=country)) +
  geom_point()
ggplotly(s) # must comment out or have eval=FALSE to knit

# boxplot
ggplotly(b) # must comment out or have eval=FALSE to knit
```




```{r, results='hide', message=FALSE, warning=FALSE}
library(tmap) # install.packages('tmap')

# load world spatial polygons
data(World)

# inspect values in World
World@data %>% tbl_df()

# gapminder countries not in World. skipping for now
g %>% 
  anti_join(World@data, by=c('country'='name')) %>% 
  arrange(desc(pop))

# World countries not in gapminder. skipping for now
World@data %>% 
  anti_join(g, by=c('name'='country')) %>% 
  arrange(desc(pop_est)) %>%
  select(iso_a3, name, pop_est)

# join gapminder data to World
World@data = World@data %>%
  left_join(g, by=c('name'='country'))
```

```{r}
# make map
m = tm_shape(World) +
    tm_polygons('lifeExp', palette='RdYlGn', id='name', title='Life expectancy (years)', auto.palette.mapping=F) +
    tm_style_gray() + tm_format_World()
m
```

## Interactive

```{r}
# show interactive map
tmap_leaflet(m)
```


# References

- [ggplot2-cheatsheet-2.0.pdf](../cheatsheets/ggplot2-cheatsheet-2.0.pdf)
- [Interactive Plots and Maps - Environmental Informatics](http://ucsb-bren.github.io/env-info/wk06_widgets.html)
- [Graphs with ggplot2 - Cookbook for R](http://www.cookbook-r.com/Graphs/#graphs-with-ggplot2)
- [ggplot2 Essentials - STHDA](http://www.sthda.com/english/wiki/ggplot2-essentials)
- [NEON Working with Geospatial Data](http://neon-workwithdata.github.io/NEON-R-Spatio-Temporal-Data-and-Management-Intro/)



